---
title: "Mission 3"
author: "Chris Jenks"
date: "2024-10-01"
output: word_document
---

```{r setup, include=FALSE}
library(readxl)
knitr::opts_chunk$set(echo = TRUE)
```

# STAT E - 100 Mission 3

```{r}
# Input the dataset
housing = read_excel("C:/Users/User/Desktop/HarvardExtensionSchool/Stat E 100/Mission 3/Housing.xlsx")
```
## 1.	Generate a table of proportions with the variable airconditioning on one axis, and hotwaterheating on the other axis. Set the margins as TRUE. Paste the R code and the output from R.

```{r}
contingency_table = table(housing$airconditioning, housing$hotwaterheating)
dimnames(contingency_table) <- list(
  AirConditioning = c("No", "Yes"), 
  HotWaterHeating = c("No", "Yes")   
)
proportions_table = prop.table(contingency_table) * 100
proportions_with_margins = addmargins(proportions_table)
print(proportions_with_margins)
```
## 2.	From the table from question 1, report the value for the following:
P(no air conditioning and no hot water heating) = **64.1 % (.640366972)**

## 3.	Generate a table of counts that has furnishingstatus on one axis, and airconditioning on another axis. Set the margins as TRUE. Include the R code and output from R.
```{r}
count_table <- table(housing$furnishingstatus, housing$airconditioning)

dimnames(count_table) = list(
  FurnishingStatus = unique(housing$furnishingstatus),  
  AirConditioning = unique(housing$airconditioning)     
)
count_with_margins = addmargins(count_table)
print(count_with_margins)
```
## 4.	For the output from question 3, calculate the following:
P(has air conditioning | furnished) = **78/140 (furnished with air / total furnished) which is 55.7% (0.55714286)**

## 5.	For the output from question 3, calculate the following:
P(has air conditioning | unfurnished) = **133/177 (unfurnished with air / total unfurnished) which is 74.7 (0.7419101)**

## 6 Determine using the data set whether the probability of a house having air conditioning and whether it is furnished are independent events. Show any calculations that lead up to the decision of whether the two events are dependent or independent.

```{r}
data_matrix = matrix(
  c(78, 62, 140, 
    162, 65, 227,
    133, 45, 178, 
    373, 172, 545),
  nrow = 4,
  byrow = TRUE
)
dimnames(data_matrix) = list(
  FurnishingStatus = c("furnished", "semi-furnished", "unfurnished", "Sum"),
  AirConditioning = c("yes", "no", "Sum")
)

total_houses = data_matrix["Sum", "Sum"]
furnished_with_ac = data_matrix["furnished", "yes"]
total_furnished = data_matrix["furnished", "Sum"]
total_with_ac = data_matrix["Sum", "yes"]
p_ac = total_with_ac / total_houses
p_furnished = total_furnished / total_houses
p_ac_and_furnished = furnished_with_ac / total_houses
p_ac_times_p_furnished = p_ac * p_furnished

cat("P(A): Probability of having air conditioning =", p_ac, "\n")
cat("P(B): Probability of being furnished =", p_furnished, "\n")
cat("P(A ∩ B): Probability of having air conditioning and being furnished =", p_ac_and_furnished, "\n")
cat("P(A) * P(B) =", p_ac_times_p_furnished, "\n\n")

if (abs(p_ac_and_furnished - p_ac_times_p_furnished) < 1e-6) {
  cat("The events are independent.\n")
} else {
  cat("The events are dependent.\n")
}
```

```{r}
cancer_exposure_table = matrix(
  c(354, 143, 293, 511),
  nrow = 2,
  byrow = TRUE
)

dimnames(cancer_exposure_table) = list(
  Exposure = c("Exposed", "Not_Exposed"),
  Cancer = c("Cancer", "No_Cancer")
)
print(cancer_exposure_table)
```
```{r}
exposed_with_cancer = cancer_exposure_table["Exposed", "Cancer"]
exposed_without_cancer = cancer_exposure_table["Exposed", "No_Cancer"]
not_exposed_with_cancer = cancer_exposure_table["Not_Exposed", "Cancer"]
not_exposed_without_cancer = cancer_exposure_table["Not_Exposed", "No_Cancer"]
total_exposed = exposed_with_cancer + exposed_without_cancer
total_not_exposed = not_exposed_with_cancer + not_exposed_without_cancer
p_cancer_given_exposed = exposed_with_cancer / total_exposed
p_cancer_given_not_exposed = not_exposed_with_cancer / total_not_exposed
relative_risk = p_cancer_given_exposed / p_cancer_given_not_exposed
cat("P(Cancer | Exposed) =", p_cancer_given_exposed, "\n")
cat("P(Cancer | Not Exposed) =", p_cancer_given_not_exposed, "\n")
cat("Relative Risk (RR) =", relative_risk, "\n")
```
## 7. Calculate the following probabilities:
### P(pass second test and pass first test) = **0.56**
### P(pass second test and fail first test) = **0.18**
### P(fail second test and pass first test) = **0.14**
### P(fail second test and fail first test) = **0.12**
### What do the four probabilities above add up to? **1**

```{r}
p_pass_first = 0.7
p_fail_first = 0.3

p_pass_second_given_pass_first = 0.8
p_fail_second_given_pass_first = 0.2

p_pass_second_given_fail_first = 0.6
p_fail_second_given_fail_first = 0.4

p_pass_second_and_pass_first = p_pass_first * p_pass_second_given_pass_first
p_pass_second_and_fail_first = p_fail_first * p_pass_second_given_fail_first
p_fail_second_and_pass_first = p_pass_first * p_fail_second_given_pass_first
p_fail_second_and_fail_first = p_fail_first * p_fail_second_given_fail_first

cat("P(pass second test and pass first test) =", p_pass_second_and_pass_first, "\n")
cat("P(pass second test and fail first test) =", p_pass_second_and_fail_first, "\n")
cat("P(fail second test and pass first test) =", p_fail_second_and_pass_first, "\n")
cat("P(fail second test and fail first test) =", p_fail_second_and_fail_first, "\n")

total_probability = p_pass_second_and_pass_first + p_pass_second_and_fail_first + 
                    p_fail_second_and_pass_first + p_fail_second_and_fail_first

cat("\nTotal of all probabilities =", total_probability, "\n")
```