---
title: "Mission 4"
author: "Chris Jenks"
date: "2024-10-08"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ggplot2)
library(readxl)
```

```{r}
bodyPerformance = read_excel("C:/Users/User/Desktop/HarvardExtensionSchool/Stat E 100/Mission 4/bodyPerformance.xlsx")
```
## 1. Use R to generate a scatterplot with sit_ups_counts on the X-axis (horizonal axis) and broad_jump_cm as the Y-axis (vertical axis). Include the R code and visualization from R.

```{r}
ggplot(bodyPerformance, aes(x = sit_ups_counts, y = broad_jump_cm)) +
  geom_point() +
  labs(x = "Sit-ups Count", y = "Broad Jump (cm)", title = "Scatterplot of Sit-ups Count vs Broad Jump (cm)") +
  theme_minimal()
```

## 2.	In the scatterplot generated in question 1), is the correlation positive, negative, or zero? Based on findings from the shape or pattern of the visualization, how do you know?
```{r}
cor(bodyPerformance$sit_ups_counts, bodyPerformance$broad_jump_cm, use = "complete.obs")
```
### The scatterplot correlation R = 0.7482728. It is positive. As such this is a positive correlation. Also, when evaluating the scatter plot, the plots both appear to be increasing. 

## 3.	Use R to calculate the sample covariance between sit_ups_counts and broad_jump_cm. Include the R code used to generate the value. Include the units of measurement for each variable.

```{r}
data_subset = bodyPerformance[, c("sit_ups_counts", "broad_jump_cm")]

cov_matrix = cov(data_subset, use = "complete.obs")
cov_matrix
```
## 4.	Calculate the sample standard deviation of sit_ups_counts and broad_jump_cm using R. Include the R code used to generate the values. Include the units of measurement for each variable
```{R}
std_dev_sit_ups = sd(bodyPerformance$sit_ups_counts, na.rm = TRUE)
std_dev_sit_ups

std_dev_broad_jump = sd(bodyPerformance$broad_jump_cm, na.rm = TRUE)
std_dev_broad_jump
```
### The standard deviation for sit ups is 14.2767. The standard deviation for broad jump is 39.868. 

## 5.	Calculate the value of the sample correlation coefficient through the formula below. Use the sample standard deviation of sit_ups_counts, the sample standard deviation of broad_jump_cm, and the sample covariance between sit_ups_counts and broad_jump_cm. Indicate a) whether the sample correlation coefficient is strong, weak, or non-existent using the rubric below this question, b) whether it is positive, negative, or near zero.
```{r}
cov_value = cov(bodyPerformance$sit_ups_counts, bodyPerformance$broad_jump_cm, use = "complete.obs")

std_dev_sit_ups = sd(bodyPerformance$sit_ups_counts, na.rm = TRUE)
std_dev_broad_jump = sd(bodyPerformance$broad_jump_cm, na.rm = TRUE)

correlation_coefficient = cov_value / (std_dev_sit_ups * std_dev_broad_jump)
correlation_coefficient
```
### The correlation coefficient is 0.7482728. This is a strong positive correlation coefficient. 

## 6.	Use R to generate the sample correlation coefficient value for sit_ups_counts and broad_jump_cm using the cor() function. Include the R code used. The value should match the value from question 5.
```{r}
correlation_value = cor(bodyPerformance$sit_ups_counts, bodyPerformance$broad_jump_cm, use = "complete.obs")
correlation_value
```
### The answer does indeed match, 0.7482728

